---
title: "Ecopath"
output: html_notebook
---

Analisis de capas dentro de modelos Ecopath

```{r load code & libraries, message=FALSE, warning=FALSE, include=FALSE}
source("libraries.R")
source("ecopath_polygons.R")
```

Usar http://ecobase.ecopath.org/#docs para coordenadas

```{r}
#To obtain the list of available model
h=basicTextGatherer()
curlPerform(url = 'http://sirs.agrocampus-ouest.fr/EcoBase/php/webser/soap-client_3.php',writefunction=h$update)

data<-xmlTreeParse(h$value(),useInternalNodes=TRUE)
liste_mod<-ldply(xmlToList(data),data.frame)

#liste_mod contains a list and decription of available models in EcoBase        


liste_mod %>% 
  dplyr::select(model.geographic_extent)

liste_mod %>% 
  write_csv("datos_ecobase.csv")

```

Study area map

```{r area map}

coords.ecopath <- read_xlsx("JCPA_Inputs Ecopaths coordinates_rev.xlsx", sheet = 1)

coords.ecopath %>% 
  dplyr::select(Ref,Year, Location, vertice, lat_rev, lon_rev) %>% 
  dplyr::rename(Zona = Location, lat= lat_rev, lon = lon_rev) %>% 
  write_csv("coordenadas_modelos_ecopath.csv")

locations.ecopath <- coords.ecopath %>% 
  distinct(Location) %>% 
  pull(Location)


lapply(locations.ecopath, make_polygons, coords.ecopath)


#shape.plot <- st_read("North and central GC.shp")

#simbolos http://www.sthda.com/english/wiki/r-plot-pch-symbols-the-different-point-shapes-available-in-r

shape.list <- do.call(rbind,lapply(list.files(path=paste0(here(),"/shapedir"), pattern = "shp$", full.names = TRUE), sf::read_sf)) %>% 
    mutate(location=if_else(location=="North and central GC","NC Gulf of California",if_else(location=="Southeastern Gulf of California","SE Gulf of California", 
                                                                                         if_else(location=="Ulloa model","Ulloa", location)))) %>% 
  filter(!location %in% c("Sinaloa benthic","Benthic GC (Sonora)","Northern GC")) %>% 
      mutate(model=factor(location, levels = c("NC Gulf of California","Isla Natividad","Bahía de Tortugas","Ulloa","La Paz","Bahía Magdalena","SE Gulf of California","Cabo Pulmo","Isla Isabel","Islas Marietas","Chamela", "Isla del Coco","Darwin and Wolf")))

eco.crs <- st_crs(shape.list)

eco.point <- shape.list %>% 
  st_transform(.,6362) %>% 
  st_centroid() %>% 
  st_transform(.,eco.crs) %>% 
   mutate(lat = unlist(map(.$geometry,2)),
           long = unlist(map(.$geometry,1)))

pacific.lme <- c("California Current","Gulf of California", "Pacific Central-American Coastal")

lme.shp <- st_read("LME66.shp") %>% 
  filter(LME_NAME %in% pacific.lme)

lme.shp.centroid <- lme.shp %>% 
  st_transform(.,6362) %>% 
  st_centroid() %>% 
  st_transform(.,eco.crs) %>% 
   mutate(ID=1:nrow(.)) %>% 
   mutate(lat = unlist(map(.$geometry,2)),
           long = unlist(map(.$geometry,1)))

lme.shp.wgs <-  st_transform(lme.shp, eco.crs)

lme.pol <- cbind(lme.shp.wgs, st_coordinates(lme.shp.centroid))

#plot points
#large map, LMEs labeled

countries <- ne_countries(10,
  "countries",
  returnclass = "sf"
) %>% 
  filter(region_un=="Americas")

countries.pol <- cbind(countries, st_coordinates(st_centroid(countries)))


point.lme <- st_join(eco.point, lme.pol, join = st_within) %>% 
  mutate(LME_NAME= if_else(model=="Cabo Pulmo", "Gulf of California", LME_NAME))

point.lme %>% 
  st_drop_geometry() %>% 
  write_csv(here("lme_model_points.csv"))

#col.pal <- brewer.pal(8, "Set2")[c(1,2,4)]

get_palette <- function(n) {
	require(RColorBrewer)
	colors <- brewer.pal.info[brewer.pal.info$colorblind == TRUE, ]
	col_vec <- unlist(mapply(brewer.pal, colors$maxcolors, rownames(colors)))
	set.seed(3478)
	return(sample(col_vec, n))
}

col.pal <- get_palette(3)

col.pal <- c("#6BAED6","#FDB863", "#7F2704")

mapa.lme <- ggplot(data = countries) +
  geom_sf() +  
  geom_sf(data = lme.pol, aes(fill=LME_NAME)) +
  scale_fill_manual(values = col.pal, labels=c("Calif. Current", "Gulf of Calif.", "Pac. Central Amer. Coastal"), name = "LME")+
  geom_text(aes(x = -103, y = 25, label = "Mexico"), stat = "unique", size = 6) +  
  geom_sf(data = eco.point) +
  #-118.33,-2.81,-76.2,36.31
    coord_sf(xlim = c(-118.33, -76.2), ylim = c(-2.81, 36.31), expand = FALSE) +
  xlab("Longitude") + ylab("Latitude") +
  theme(panel.grid.major = element_line(color = gray(0.4), linetype = "dashed", 
                                        size = 0.1), panel.background = element_rect(fill = "white")) +
  annotation_scale(location = "bl", width_hint = 0.4) +
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
                         style = north_arrow_fancy_orienteering) +
  theme(legend.position="bottom", axis.text = element_text(size = 11), axis.title = element_text(size = 13), legend.text = element_text(size = 13), legend.title = element_text(size = 14))

ggsave(here("map_lme.png"), mapa.lme, width = 10, height = 10)
  
eco.north <- eco.point %>% 
  filter(!model %in% c("Isla del Coco","Darwin and Wolf")) %>% 
  mutate(map_index = c(3,6,7,11,9,2,10,5,1,8,4))

mapa.secc1 <- ggplot(data = countries) +
  geom_sf() +  
  geom_sf(data = lme.pol, aes(fill=LME_NAME)) +
  geom_text(aes(x = -107, y = 28, label = "Mexico"), stat = "unique", size = 5) +  
  scale_fill_manual(values = col.pal)+
  geom_sf(data = eco.north) + 
  geom_sf(data = shape.list, fill = NA, color = gray(.7)) +
  #-118.83,17.6,-100.72,32.19
  coord_sf(xlim = c(-118.83, -100.72), ylim = c(17.6, 32.19), expand = FALSE) +
  xlab("") + ylab("") +
  theme(panel.grid.major = element_line(color = gray(0.4), linetype = "dashed", 
                                        size = 0.1), panel.background = element_rect(fill = "white")) +
  annotation_scale(location = "bl", width_hint = 0.4) +
  geom_text_repel(data = eco.north, aes(x = long, y = lat, label = map_index), 
        fontface = "plain", nudge_x = c(-1.5, -1.5, 1, 1, 1,-2,-2,1,2,-2,-1), nudge_y = c(-0.75, -0.25, -0.5,0,0,0.5,0,0,0.5,-0.5,0), size = 5) +
theme(legend.position="bottom", axis.text = element_text(size = 11), axis.title = element_text(size = 13), legend.text = element_text(size = 11), legend.title = element_text(size = 13)) +
    theme(legend.position="none") 


ggsave(here("map_north.png"), mapa.secc1, width = 5, height = 5)

eco.south <- eco.point %>% 
  filter(model %in% c("Isla del Coco","Darwin and Wolf")) %>% 
  mutate(map_index = c(13,12))


mapa.secc2 <- ggplot(data = countries) +
  geom_sf() +  
  geom_sf(data = lme.pol, aes(fill=LME_NAME)) +
  geom_text(aes(x = -79.2, y = -1, label = "Ecuador"), stat = "unique", size = 5) +  
  scale_fill_manual(values = col.pal, name = "LME")+
  geom_sf(data = eco.south) + 
  geom_sf(data = shape.list, fill = NA, color = gray(.7)) +
  #-94.22,-6.07,-76.11,10.01
  coord_sf(xlim = c(-94.22, -76.11), ylim = c(-6.07, 10.01), expand = FALSE) +
  xlab("") + ylab("") +
  theme(panel.grid.major = element_line(color = gray(0.4), linetype = "dashed", 
                                        size = 0.1), panel.background = element_rect(fill = "white")) +
  annotation_scale(location = "bl", width_hint = 0.4) +
  theme(legend.position="none") +
  geom_text_repel(data = eco.south, aes(x = long, y = lat, label = map_index), 
        fontface = "plain", nudge_x = c(1, -1.5), nudge_y = c(1, 0.25), size = 5) +
  theme(legend.position="bottom", axis.text = element_text(size = 11), axis.title = element_text(size = 13), legend.text = element_text(size = 11), legend.title = element_text(size = 13)) + 
     theme(legend.position="none") 

ggsave(here("map_south.png"), mapa.secc2, width = 5, height = 5)

#map.grid <- mapa.lme | mapa.secc1 | mapa.secc2

map.grid <- mapa.lme + (mapa.secc1 / mapa.secc2) + plot_layout(widths = c(3, 2))

ggsave(here("map_ecopath.png"), map.grid, width = 10, height = 6)

```


```{r}

wgs.crs <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
 
base.raster <- raster(paste0(here(),"/rasters/pp_mean.asc"))
values(base.raster)[values(base.raster) > 0] = 0

shape.list <- do.call(rbind,lapply(list.files(path=paste0(here(),"/shapedir"), pattern = "shp$", full.names = TRUE), sf::read_sf)) %>% 
  mutate(location=if_else(location=="North and central GC","NC Gulf of California",if_else(location=="Southeastern Gulf of California","SE Gulf of California", 
                                                                                         if_else(location=="Ulloa model","Ulloa", location)))) %>% 
  filter(!location %in% c("Sinaloa benthic","Benthic GC (Sonora)","Northern GC")) %>% 
      mutate(model=factor(location, levels = c("NC Gulf of California","Isla Natividad","Bahía de Tortugas","Ulloa","La Paz","Bahía Magdalena","SE Gulf of California","Cabo Pulmo","Isla Isabel","Islas Marietas","Chamela", "Isla del Coco","Darwin and Wolf")))

extent.ecopath <- extent(base.raster)

areapolygon <- paste0("POLYGON((",extent.ecopath@xmin," ", extent.ecopath@ymax,",", extent.ecopath@xmax," ",extent.ecopath@ymax, ",", extent.ecopath@xmax," ", extent.ecopath@ymin, ",", extent.ecopath@xmin, " ", extent.ecopath@ymin,",", extent.ecopath@xmin," ", extent.ecopath@ymax,"))")

  obis.data.table  <- tryCatch(robis::occurrence(geometry =areapolygon, fields=c("species","decimalLongitude", "decimalLatitude"),verbose=TRUE),error=function(e) as.character()) 
   
richness.table <- obis.data.table %>% 
  distinct(species, decimalLatitude, decimalLongitude)

write_csv(obis.data.table, "obis_records.csv")

coordinates(obis.data.table) <- c("decimalLongitude", "decimalLatitude")  # set spatial coordinates
proj4string(obis.data.table) <- wgs.crs  # define projection system of our data
 
richness.raster <- rasterize(obis.data.table, base.raster, fun='count')
proj4string(richness.raster) <- wgs.crs  # define projection system of our 

rs1 <- calc(richness.raster, sum) %>% 
  rescale0to1()

proj4string(rs1) <- wgs.crs  # define projection system of our 

writeRaster(rs1, paste0(here(),"/rasters/richness.asc"), format="ascii", overwrite=TRUE)
    

rs2 <- calc(richness.raster, mean) %>% 
  rescale0to1()

proj4string(rs2) <- wgs.crs  # define projection system of our 

writeRaster(rs2, paste0(here(),"/rasters/richness_mean.asc"), format="ascii", overwrite=TRUE)

```

rasters para valores

informacion de la base de perturbaciones humanas halpern 2013: 
efecto acumulativo, 
pesca artesanal
ocean pollution
shipping
productividad primaria
riqueza de peces e invertebrados
promedio, varianza y N pixeles con datos espaciales

```{r}

wgs.crs <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

mol.crs <- CRS("+proj=moll +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs")

base.raster <- raster(paste0(here(),"/rasters/pp_mean.asc"))
values(base.raster)[values(base.raster) > 0] = 0
crs(base.raster) <- wgs.crs

# productividad primaria
# prod.raster.atlantic <- raster(paste0(here(),"/rasters/pp_mean_atlantic.asc"))
# prod.raster.pacific <- raster(paste0(here(),"/rasters/pp_mean_pacific.asc"))
# 
#prod.raster <- raster::merge(prod.raster.atlantic, prod.raster.pacific) %>%   rescale0to1()
# 
# writeRaster(prod.raster, paste0(here(),"/rasters/pp_mean.asc"), format="ascii")
  

impact.raster <- raster(paste0(here(),"/rasters/global_cumul_impact_2013_all_layers.tif"))
impact.raster.pr <- projectRaster(impact.raster, crs = wgs.crs)

impact.raster.rs <- resample(impact.raster.pr, base.raster, method = 'bilinear') # resample output
impact.raster.sc <- rescale0to1(impact.raster.rs)


writeRaster(impact.raster.sc, paste0(here(),"/rasters/impact.asc"), format="ascii", overwrite=TRUE)


artisanal.raster <- raster(paste0(here(),"/rasters/artisanal_fishing.tif")) 
artisanal.raster.pr <-projectRaster(artisanal.raster, crs = wgs.crs)

artisanal.raster.rs <- resample(artisanal.raster.pr, base.raster, method = 'bilinear') # resample output

artisanal.raster.sc <- rescale0to1(artisanal.raster.rs)
  
writeRaster(artisanal.raster.sc, paste0(here(),"/rasters/artisanal.asc"), format="ascii", overwrite=TRUE)


shipping.raster <- raster(paste0(here(),"/rasters/shipping.tif")) shipping.raster.pr <-projectRaster(shipping.raster, crs = wgs.crs)

shipping.raster.rs <- resample(shipping.raster.pr, base.raster, method = 'bilinear') # resample output
shipping.raster.sc <- rescale0to1(shipping.raster.rs)

writeRaster(shipping.raster.sc, paste0(here(),"/rasters/shipping.asc"), format="ascii", overwrite=TRUE)

#contaminacion
pollution.raster <- raster(paste0(here(),"/rasters/ocean_pollution_combo.tif")) 
pollution.raster.pr <-projectRaster(pollution.raster, crs = wgs.crs)

pollution.raster.rs <- resample(pollution.raster.pr, base.raster, method = 'bilinear') # resample output
pollution.raster.sc <- rescale0to1(pollution.raster.rs)

writeRaster(pollution.raster.sc, paste0(here(),"/rasters/pollution.asc"), format="ascii", overwrite=TRUE)


```


Extraer valores por poligono
```{r}

#productividad
prod.raster.calc <- raster(paste0(here(),"/rasters/pp_mean.asc"))
crs(prod.raster.calc) <- wgs.crs

#riqueza
richness.raster.calc <- raster(paste0(here(),"/rasters/richness.asc"))
crs(richness.raster.calc) <- wgs.crs

#pollution
pollution.raster.calc <- raster(paste0(here(),"/rasters/pollution.asc"))
crs(pollution.raster.calc) <- wgs.crs

#artisanal fisheries
artisanal.raster.calc <- raster(paste0(here(),"/rasters/artisanal.asc"))
crs(artisanal.raster.calc) <- wgs.crs

#impact
impact.raster.calc <- raster(paste0(here(),"/rasters/impact.asc"))
crs(impact.raster.calc) <- wgs.crs

#shipping
shipping.raster.calc <- raster(paste0(here(),"/rasters/shipping.asc"))
crs(shipping.raster.calc) <- wgs.crs


shape.list <- do.call(rbind,lapply(list.files(path=paste0(here(),"/shapedir"), pattern = "shp$", full.names = TRUE), sf::read_sf)) 


richness.mean <- exact_extract(richness.raster.calc, shape.list , 'mean', stack_apply = TRUE)
prod.mean <- exact_extract(prod.raster.calc, shape.list , 'mean')
artisanal.mean <- exact_extract(artisanal.raster.calc, shape.list , 'mean')
pollution.mean <- exact_extract(pollution.raster.calc, shape.list , 'mean')
shipping.mean <- exact_extract(shipping.raster.calc, shape.list , 'mean')
impact.mean <- exact_extract(impact.raster.calc, shape.list , 'mean')


richness.val <- exact_extract(richness.raster.calc, shape.list, 'variance')
prod.val <- exact_extract(prod.raster.calc, shape.list , 'variance')
artisanal.val <- exact_extract(artisanal.raster.calc, shape.list , 'variance')
pollution.val <- exact_extract(pollution.raster.calc, shape.list , 'variance')
shipping.val <- exact_extract(shipping.raster.calc, shape.list , 'variance')
impact.val <- exact_extract(impact.raster.calc, shape.list , 'variance')


richness.count <- exact_extract(richness.raster.calc, shape.list, 'count')
prod.count <- exact_extract(prod.raster.calc, shape.list , 'count')
artisanal.count <- exact_extract(artisanal.raster.calc, shape.list , 'count')
pollution.count <- exact_extract(pollution.raster.calc, shape.list , 'count')
shipping.count <- exact_extract(shipping.raster.calc, shape.list , 'count')
impact.count <- exact_extract(impact.raster.calc, shape.list , 'count')

#make table of data
shape.names <- list.files(path=paste0(here(),"/shapedir"), pattern = "shp$", full.names = TRUE) %>% 
  gsub("/home/atlantis/ecopath/shapedir/","",.) %>% 
  gsub(".shp","",.)

summary.index <- tibble(modelo = shape.names, riqueza_media = richness.mean, productividad_media = prod.mean, artesanal_media = artisanal.mean, contaminacion_media = pollution.mean, trafico_barcos_media= shipping.mean, impacto_media = impact.mean, riqueza_var = richness.val, productividad_var = prod.val, artesanal_var = artisanal.val, contaminacion_var = pollution.val, trafico_barcos_var= shipping.val, impacto_var = impact.val, riqueza_conteo = richness.count, productividad_conteo = prod.count, artesanal_conteo = artisanal.count, contaminacion_conteo = pollution.count, trafico_barcos_conteo= shipping.count, impacto_conteo = impact.count)

write_csv(summary.index, "resumen_indices_ecopath.csv")

```

Summary models and ranked correlations

```{r summary models}

summary.models <- read_xlsx("Summary models PiO.xlsx", sheet = 1) %>% 
  pivot_longer(cols = -Variable, names_to = "model",values_to = "value") %>% 
  pivot_wider(id_cols=model, names_from = Variable, values_from = "value") %>% 
  dplyr::select(!ends_with("_conteo")) %>% 
   dplyr::select(!ends_with("_var"))

#Ascendency, Overhead and the Overhead/ Capacity ratio as measures of resilience and the Connectance and Omnivory indices as measures of resistance, a

summary.models.plot <- read_xlsx("Summary models PiO.xlsx", sheet = 1) %>% 
  pivot_longer(cols = -Variable, names_to = "model",values_to = "value") %>% 
  mutate(Variable=if_else(Variable=="riqueza_media","Richness",
                          if_else(Variable=="productividad_media","Productivity",
                                  if_else(Variable=="trafico_barcos_media","Boat traffic",
                                          if_else(Variable== "artesal_media", "Artisanal fisheries",
                                                  if_else(Variable=="contamicion_media","Pollution",
                                                          if_else(Variable== "impacto_media","Cumulative impact",Variable)))))))

normalit<-function(m){
   (m - min(m))/(max(m)-min(m))
}

lme.points <- read_csv(here("lme_model_points.csv")) %>% 
        mutate(model=if_else(model=="Diaz-Uribe GC","NC Gulf of California",
                             if_else(model=="Southeastern","SE Gulf of California", 
                                     if_else(model=="Isla Natividad", "I. Natividad", 
                                             if_else(model=="Bahía Magdalena", "B. Magdalena", model))))) %>% 
  mutate(LME_NAME = if_else(model == "Isla del Coco" | model == "Darwin and Wolf", "Oceanic islands", LME_NAME))


model.plot.data <- summary.models.plot %>% 
      filter(Variable %in% c("Richness","Productivity","Boat traffic","Artisanal fisheries","Pollution","Cumulative impact")) %>% 
      mutate(Variable = as.factor(Variable)) %>% 
      mutate(value = if_else(is.na(value),0,value)) %>% 
      group_by(Variable) %>% 
      resca(value, new_min = 0, new_max = 1) %>% 
      mutate(model=if_else(model=="Diaz-Uribe GC","NC Gulf of California",if_else(model=="Southeastern","SE Gulf of California", model))) %>% 
  left_join(lme.points, by="model") %>% 
        mutate(model=factor(model, levels = c("NC Gulf of California","I. Natividad","Bahía de Tortugas","Ulloa","La Paz","B. Magdalena","SE Gulf of California",
                                      "Cabo Pulmo","Isla Isabel","Islas Marietas","Chamela", "Isla del Coco","Darwin and Wolf")))


model.plot <- model.plot.data %>% 
      ggplot(aes(x=Variable, y=value_res, group=model, fill = LME_NAME)) + 
      geom_bar(stat = "identity")+
  #geom_line() +
  scale_fill_brewer(palette ="Set2", name = "LME")+
  facet_wrap(~ model, scales="free_y")+
      theme_classic()+
      ylab("Value") + 
      xlab("Model") +
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_text(angle = 90))
   
ggsave("model_variables.png", model.plot, device="png",width=9,height = 7,dpi=350)
    


#Ranked cross-correlations

#Ascendency, Overhead and the Overhead/ Capacity ratio as measures of resilience and the Connectance and Omnivory indices as measures of resistance, 

ranked.cross <- summary.models %>% 
  mutate_if(is.numeric, replace_na, '0') %>% 
  dplyr::select(ends_with("media")) %>% 
  corr_cross(., # name of dataset
  max_pvalue = 0.01, # display only significant correlations (at 5% level)
  top = 30 # display top 10 couples of variables (by correlation coefficient)
)

ggsave("rank_correlations_plot.png", ranked.cross, device="png",width=10,height=9, dpi=300)


#calculate multicollinearity
#http://www.sthda.com/english/articles/39-regression-model-diagnostics/160-multicollinearity-essentials-and-vif-in-r/


summary.models.lm <- summary.models %>% 
   dplyr::select(ends_with("media"), Ascendency) %>% 
  dplyr::rename(Productivity = productividad_media, Richness=riqueza_media, `Artisanal fisheries`= artesal_media, "Boat traffic"=trafico_barcos_media, Pollution=contamicion_media, "Cumulative impact" = "impacto_media")


model1 <- lm(-Ascendency ~ ., data = summary.models.lm, na.action = na.exclude)

vif.res <- car::vif(model1) %>% 
  as.data.frame() %>% 
  as_tibble(rownames = "variable") %>% 
  setNames(c("variable","VIF")) 


vif.res$variable_factor <- factor(vif.res$variable, c("Productivity","Richness","Artisanal fisheries","Boat traffic","Pollution","Cumulative impact")) 

col.pal2 <- c(redmonder.pal(9,"sPBIBu")[2:9],redmonder.pal(9,"sPBIPu")[4:7],redmonder.pal(9,"sPBIRdPu")[2:5])

vif.plot <- ggplot(aes(x=variable,y=VIF, fill=variable), data=vif.res) +
  geom_bar(stat="identity") +
  scale_fill_manual(values=col.pal2)+
   theme(axis.text.x = element_blank())+
  labs(x="Variable")+
  geom_hline(yintercept=3)

ggsave("vif_multicolinearity_plot.png", vif.plot, device="png",width=10,height=8, dpi=200)

write_csv(vif.res,"VIF_values.csv")






```

#all variables
```{r}

#June 2022 TST, NPP, PP/B, B/TST, overhead y A/C.

model.indicators <- c("Total system throughput", "Calculated total net primary production", "Total primary production/total biomass", "Total biomass/total throughput", "Overhead", "A/C")

new.model.indicators <- c("total_system_throughput", "primary_production", "primary_prod_biomass", "tot_biomass_tot_throughput", "overhead", "A_C")
  
  
summary.ind <- read_xlsx("Summary models PiO.xlsx", sheet = 1) %>% 
   pivot_longer(cols = -Variable, names_to = "model",values_to = "value") %>% 
  dplyr::filter(Variable %in% model.indicators) %>% 
  mutate(value = if_else(is.na(value),0,value)) %>% 
  group_by(Variable) %>% 
 # resca(value, new_min = 0, new_max = 1) %>% 
 # dplyr::select(-value) %>% 
  pivot_wider(names_from = Variable, values_from = value) %>% 
  mutate(rec_index=1:nrow(.)) %>% 
  rename_with(~ new.model.indicators, all_of(model.indicators))


summary.vars <- read_xlsx("Summary models PiO.xlsx", sheet = 1) %>% 
   pivot_longer(cols = -Variable, names_to = "model",values_to = "value") %>% 
  dplyr::filter(grepl("media",Variable)) %>% 
  mutate(value = if_else(is.na(value),0,value)) %>% 
  group_by(Variable) %>% 
  resca(value, new_min = 0, new_max = 1) %>% 
  dplyr::select(-value) %>% 
  pivot_wider(names_from = Variable, values_from = value_res) %>% 
  dplyr::rename("Richness"=riqueza_media,"Productivity"=productividad_media,"Artisanal fisheries"=artesal_media,"Pollution"=contamicion_media,"Boat traffic"=trafico_barcos_media,"Cumulative impact"=impacto_media) %>% 
  dplyr::select(-model)

data.summary <- cbind(summary.ind, summary.vars)


get_glm <- function(this.variable, data.summary){
  
  print(this.variable)
  
  pc.glm <-  glm(formula = substitute(i ~ Productivity + `Artisanal fisheries` + `Cumulative impact`, list(i = as.name(this.variable))), family= gaussian, data=data.summary)
  
  print(summary(pc.glm))
  
  saveRDS(pc.glm,here("model_results",paste0("glm_",this.variable, ".rds")))
  
  
return(pc.glm)

} 

glm.models <- lapply(new.model.indicators, get_glm, data.summary)

```


#PCA

```{r}
#PCA on variables 

#net primary production, PP/B, Biomass/ Total system throughput, connectance index, ascendency, overhead and development capacity

model.indicators <- c("Calculated total net primary production", "Total primary production/total biomass", "Total biomass/total throughput", "Connectance Index", "Ascendency","Overhead", "Capacity")

new.model.indicators <- c("primary_production", "primary_prod_biomass", "tot_biomass_tot_throughput", "connectance", "ascendency","overhead", "capacity")
  
  
summary.ind <- read_xlsx("Summary models PiO.xlsx", sheet = 1) %>% 
   pivot_longer(cols = -Variable, names_to = "model",values_to = "value") %>% 
  dplyr::filter(Variable %in% model.indicators) %>% 
  mutate(value = if_else(is.na(value),0,value)) %>% 
  group_by(Variable) %>% 
 # resca(value, new_min = 0, new_max = 1) %>% 
 # dplyr::select(-value) %>% 
  pivot_wider(names_from = Variable, values_from = value) %>% 
  mutate(rec_index=1:nrow(.)) %>% 
  rename_with(~ new.model.indicators, all_of(model.indicators))


summary.vars <- read_xlsx("Summary models PiO.xlsx", sheet = 1) %>% 
   pivot_longer(cols = -Variable, names_to = "model",values_to = "value") %>% 
  dplyr::filter(grepl("media",Variable)) %>% 
  mutate(value = if_else(is.na(value),0,value)) %>% 
  group_by(Variable) %>% 
  resca(value, new_min = 0, new_max = 1) %>% 
  dplyr::select(-value) %>% 
  pivot_wider(names_from = Variable, values_from = value_res) %>% 
  dplyr::rename("Richness"=riqueza_media,"Productivity"=productividad_media,"Artisanal fisheries"=artesal_media,"Pollution"=contamicion_media,"Boat traffic"=trafico_barcos_media,"Cumulative impact"=impacto_media) %>% 
  dplyr::select(-model)


stand.environ.index <- vegan::decostand(summary.vars, "standardize") %>% 
  mutate(rec_index=1:nrow(.)) %>% 
  na.omit

stand.environ <- vegan::decostand(summary.vars, "standardize") %>% 
  na.omit 

 # Do the PCA 
#To perform principal component analysis using the correlation matrix using the prcomp() function, set the scale argument to TRUE.
#This approach is recommended for data that is measured in different units or has wide variances.

res.pca <- prcomp(stand.environ,center=T,scale=T)

res.cov <- cov(stand.environ)
res.eigen <- eigen(res.cov)

PVE <- res.eigen$values / sum(res.eigen$values)

#choose how many dimensions you need
stand.environ.paran <- paran(stand.environ, graph = TRUE)

parallel.analysis <- tibble(AdjustedEv=stand.environ.paran$AdjEv, RandomEv= stand.environ.paran$RndEv, UnadjustedEv=stand.environ.paran$Ev, PVE=PVE) %>% 
  mutate(components=1:nrow(.)) %>% 
  pivot_longer(cols=AdjustedEv:PVE, names_to = "parameter", values_to="eigenvalue") 

col.pal1 <- c(redmonder.pal(8,"qMSOSlp"))
col.pal1 <- col.pal1[c(1,3,4)]

paralell.plot <- parallel.analysis %>% 
  filter(parameter!="PVE") %>% 
  filter(components < 11) %>% 
  mutate(components = as.factor(components)) %>% 
  ggplot(aes(x=components,y=eigenvalue, colour=parameter, group=parameter))+
  geom_line()+
  scale_colour_manual(values=col.pal1, name="Parameter", labels=c("Adjusted Ev","Random Ev","Unadjusted Ev")) +
 # geom_hline(yintercept=1) +
  theme_minimal() +
  xlab("Dimensions") +
  ylab("Eigenvalue") +
  ggtitle("Parallel analysis")


#scree plot
#the plot of eigenvalues ordered from largest to the smallest. The number of component is determined at the point, beyond which the remaining eigenvalues are all relatively small and of comparable size (Jollife 2002, Peres-Neto, Jackson, and Somers (2005)).
scree.plot <- fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50)) +
  theme_minimal()

pca.test.plot <- grid.arrange(paralell.plot,scree.plot, nrow=2, ncol=1)

ggsave("dimensions_plot_pca.png", pca.test.plot, device="png",width=8,height = 6,dpi=350)



#graph of variables
var.pca.plot <- fviz_pca_var(res.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

ggsave("var_plot_pca.png", var.pca.plot, device="png",width=10,height = 8,dpi=350)


# #graph of individual data points
# fviz_pca_ind(res.pca,
#              col.ind = "cos2", # Color by the quality of representation
#              gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
#              repel = TRUE     # Avoid text overlapping
#              )

#pca.plot <- autoplot(res.pca, data = stand.environ)
#pca.plot

# Eigenvalues
#http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/112-pca-principal-component-analysis-essentials/
# examine the eigenvalues to determine the number of principal components to be considered. 

#An eigenvalue > 1 indicates that PCs account for more variance than accounted by one of the original variables in standardized data. This is commonly used as a cutoff point for which PCs are retained. This holds true only when the data are standardized.

#You can also limit the number of component to that number that accounts for a certain fraction of the total variance. For example, if you are satisfied with 70% of the total variance explained then use the number of components to achieve that.

eig.val <- get_eigenvalue(res.pca)
eig.val
  
# Results for Variables
res.var <- get_pca_var(res.pca)
res.var$coord          # Coordinates
res.var$contrib        # Contributions to the PCs
res.var$cos2           # Quality of representation 


#The quality of representation of the variables on factor map is called cos2 (square cosine, squared coordinates)

png(height=800, width=800, file="corr_plot_pca.png", type = "cairo")
corrplot(res.var$cos2, 
         method = "circle",
         is.corr = F,
         order = "original",
         tl.col="black")
dev.off()



# Contributions of variables to PC1
pca.comp1 <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10)+
  ggtitle("Dim 1")
pca.comp2 <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10)+
  ggtitle("Dim 2")
pca.comp3 <- fviz_contrib(res.pca, choice = "var", axes = 3, top = 10)+
  ggtitle("Dim 3")
pca.comp4 <- fviz_contrib(res.pca, choice = "var", axes = 4, top = 10)+
  ggtitle("Dim 4")
pca.comp5 <- fviz_contrib(res.pca, choice = "var", axes = 5, top = 10)+
  ggtitle("Dim 5")
pca.comp6 <- fviz_contrib(res.pca, choice = "var", axes = 6, top = 10)+
  ggtitle("Dim 6")

var.contribution <- grid.arrange(pca.comp1,pca.comp2,pca.comp3,pca.comp4,pca.comp5,pca.comp6, nrow=3, ncol = 2)

# Total cos2 of variables on Dim.1 to Dim.4
fviz_cos2(res.pca, choice = "var", axes = 1)

# A high cos2 indicates a good representation of the variable on the principal component. In this case the variable is positioned close to the circumference of the correlation circle.
# 
# A low cos2 indicates that the variable is not perfectly represented by the PCs. In this case the variable is close to the center of the circle.



#attach pc axes later selected to data
pca.axes <- res.pca$x %>% 
  as_tibble %>% 
  bind_cols(summary.ind) %>% 
  mutate(rec_index = as.factor(rec_index)) %>% 
  bind_cols(summary.vars) 
  
#GAM resources
#https://m-clark.github.io/generalized-additive-models/application.html#multiple-predictors


#gamma 1.4 to reduce overfitting

#the Variance Inflation Factor, which determines multicollinearity found values > 5 (Figure S5) for PDO, Harbour seals, steelhead salmon, pink salmon, Cherry point herring stock, and orcas. As a result, we eliminated these five variables from the final models.

#full model with no interaction terms


get_glm <- function(this.variable, pca.axes){
  
  print(this.variable)
  
  pc.glm <-  glm(formula = substitute(i ~ PC1 + PC2 + PC3, list(i = as.name(this.variable))), family= gaussian, data=pca.axes)
  
  saveRDS(pc.glm,here("model_results",paste0("pc_glm_",this.variable, ".rds")))
  
  
return(pc.glm)

} 

glm.pca.models <- lapply(new.model.indicators, get_glm, pca.axes)


```

GLM models that only use population as a proxy for human impact


```{r}

# O/C and A/C are proxies for resilience
model.indicators <- c("O/C", "A/C")

new.model.indicators <- c("O_C", "A_C")

#load model indicators
summary.ind <- read_xlsx("Summary models PiO.xlsx", sheet = 1) %>% 
   pivot_longer(cols = -Variable, names_to = "model",values_to = "value") %>% 
  dplyr::filter(Variable %in% model.indicators) %>% 
  mutate(value = if_else(is.na(value),0,value)) %>% 
  group_by(Variable) %>% 
 # resca(value, new_min = 0, new_max = 1) %>% 
 # dplyr::select(-value) %>% 
  pivot_wider(names_from = Variable, values_from = value) %>% 
  mutate(rec_index=1:nrow(.)) %>% 
  rename_with(~ new.model.indicators, all_of(model.indicators)) %>% 
  mutate(model = gsub("Bahía","Bahia",model))

summary.var <- read_csv("PIO_oceanography.csv") %>% 
  dplyr::select(model, chl_a, SST_11, POC)  


eco.data <- summary.ind %>% 
  left_join(summary.var, by = c("model")) %>% 
  mutate(model = gsub("Bahía","Bahia",model)) %>% 
  mutate(model = if_else(model=="Bahia Magdalena","B. Magdalena", model))

ranked.cross <- summary.var %>%
  dplyr::select(-model) %>% 
  mutate_if(is.numeric, replace_na, '0') %>% 
  corr_cross(., # name of dataset
  max_pvalue = 0.01, # display only significant correlations (at 5% level)
  top = 30 # display top 10 couples of variables (by correlation coefficient)
)

ggsave("rank_correlations_plot.png", ranked.cross, device="png",width=10,height=9, dpi=300)


model1 <- lm(-A_C ~ ., data = eco.data, na.action = na.exclude)

vif.res <- car::vif(model1) %>% 
  as.data.frame() %>% 
  as_tibble(rownames = "variable") %>% 
  setNames(c("variable","VIF")) 


vif.res$variable_factor <- factor(vif.res$variable, c("Productivity","Richness","Artisanal fisheries","Boat traffic","Pollution","Cumulative impact")) 

col.pal2 <- c(redmonder.pal(9,"sPBIBu")[2:9],redmonder.pal(9,"sPBIPu")[4:7],redmonder.pal(9,"sPBIRdPu")[2:5])

vif.plot <- ggplot(aes(x=variable,y=VIF, fill=variable), data=vif.res) +
  geom_bar(stat="identity") +
  scale_fill_manual(values=col.pal2)+
   theme(axis.text.x = element_blank())+
  labs(x="Variable")+
  geom_hline(yintercept=3)

ggsave("vif_multicolinearity_plot.png", vif.plot, device="png",width=10,height=8, dpi=200)

write_csv(vif.res,"VIF_values.csv")

#model polygons
shape.ecopath <- do.call(rbind,lapply(list.files(path=paste0(here(),"/shapedir"), pattern = "shp$", full.names = TRUE), sf::read_sf)) %>% 
  mutate(location = gsub("Bahía","Bahia",location)) %>% 
  mutate(location=if_else(location=="North and central GC","NC Gulf of California",
                          if_else(location=="Southeastern Gulf of California","Southeastern",
                                  if_else(location=="Ulloa model","Ulloa", 
                                          if_else(location=="Isla Natividad","I. Natividad", 
                                                  if_else(location=="Bahia Magdalena","B. Magdalena", location)))))) %>% 
  filter(!location %in% c("Sinaloa benthic","Benthic GC (Sonora)","Northern GC")) %>% 
      mutate(model=factor(location, levels = c("NC Gulf of California","I. Natividad","Bahia de Tortugas","Ulloa","La Paz","B. Magdalena","SE Gulf of California","Cabo Pulmo","Isla Isabel","Islas Marietas","Chamela", "Isla del Coco","Darwin and Wolf")))


# make 30 km buffer around model polygons

wgs.crs <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

shape.ecopath.wgs <- st_transform(shape.ecopath, wgs.crs)

shape.eco.buffer <- shape.ecopath.wgs %>% 
  st_transform(.,6362) %>%
  st_buffer(.,dist = 3000)
  
#load INEGI layer

comm.costas <- st_read(here("comunidades_costeras_30km.shp")) %>% 
  st_transform(.,wgs.crs) %>% 
    st_transform(.,6362)

#extract population within buffered model polygons

comm.in.pol <- st_join(comm.costas, shape.eco.buffer, join = st_within)

#aggregate into single column and assign to model

pob.count <- as_tibble(comm.in.pol) %>% 
  dplyr::group_by(location) %>% 
  dplyr::summarise(tot_pob=sum(POBTOT)) %>% 
  dplyr::rename(model=location) %>% 
  mutate(model = gsub("Bahía","Bahia",model)) 
# run glm with population

  
model.data <- eco.data %>% 
  left_join(pob.count, by="model") %>% 
  mutate(tot_pob = if_else(is.na(tot_pob), 0, tot_pob))
  
  
 
#Modelo 1:
#O/C ~(a + bChla + cSST + dPOC + ePobHum)
#Modelo 2:
#A/C ~(a + bChla + cSST + dPOC + ePobHum)

get_glm <- function(this.variable, data.summary){
  
  print(this.variable)
  
 # pc.glm <-  glm(formula = substitute(i ~ chl_a + SST_11 + tot_pob, list(i = as.name(this.variable))), family= gaussian, data=model.data)
  #pc.glm <-  glm(formula = substitute(i ~ chl_a + SST_11 + POC + tot_pob, list(i = as.name(this.variable))), family= gaussian, data=model.data)
  pc.glm <-  glm(formula = substitute(i ~ chl_a + SST_11 + tot_pob, list(i = as.name(this.variable))), family= gaussian, data=model.data)
   

  print(summary(pc.glm))
  
  saveRDS(pc.glm,here("model_results",paste0("glm_",this.variable, ".rds")))
  
  
return(pc.glm)

} 

glm.models <- lapply(new.model.indicators, get_glm, model.data)



```



Resources used
Rescaling
https://stackoverflow.com/questions/32278825/how-to-change-the-resolution-of-a-raster-layer-in-r
https://search.r-project.org/CRAN/refmans/climateStability/html/rescale0to1.html
Metadata Halpern
https://knb.ecoinformatics.org/view/doi%3A10.5063%2FF19C6VN5
Mollweide projection
https://epsg.io/54009
Merge
https://www.rdocumentation.org/packages/raster/versions/3.4-13/topics/merge
Extract
https://isciences.gitlab.io/exactextractr/reference/exact_extract.html